This chapter reviews the state of energy system optimization modeling (ESOM) tools
and tools for evaluating nuclear hybrid energy systems (NHES). The role of \glspl{esom}
in energy and environmental policymaking is also discussed.

Energy systems are spatially, temporally, and topologically complex ‘machines’
that operate cohesively to supply energy where it is needed at the exact moment
it is needed. Spatial complexity arises from various interconnections that enable
energy transmission from one region to another. This includes transmission
lines that move electricity from one part of the United States to another or
pipelines moving natural gas. Regions can be large multistate entities like the
PJM Interconnection, smaller communities, or even individual homes. Temporal
complexity arises from the time dependence of energy demand and the availability of
variable and intermittent energy sources. Energy demand and renewable energy
production
exhibit a variety of seasonalities from annual, monthly, even daily trends.
At the minute to hourly level, there is also some stochasticity caused either by
human behavior or weather conditions.
Examining energy systems at each of these temporal resolutions is vital for
generating robust energy policy and guaranteeing reliable energy delivery. Finally,
topological complexity refers to the number of interacting technologies that can
produce and move energy and the number of unique energy demands.
A community powered by a small solar farm and a diesel
backup generator is topologically simpler than a state with many types of energy
demands, various means of production, and dozens of individual generators. Beyond
the physical and engineering aspects of energy systems is a social context that
motivates choices at every layer of complexity. These choices reflect the priorities
of a given society or community and manifest as energy policy.

\section{Energy Policy and Climate Change Mitigation}
\label{section:energy_policy}

Energy policies serve to coordinate the different aspects of energy systems.
Essentially, energy policies need to ensure reliability, affordability, and
sustainability \cite{fattahi_systemic_2020}. A reliable grid ensures an adequate
energy supply when needed \cite{milligan_methods_2011, ramirez-meyers_how_2021,berkeley_iii_framework_2010}.
An unreliable grid experiences an unacceptable frequency of outages or losses of
load \cite{ramirez-meyers_how_2021}. The North American Electric Reliability
Corporation sets the desired loss-of-load frequency at 0.1 days/year and recommends
a \gls{prm} of 15 percent \cite{milligan_methods_2011,reimers_impact_2019}. \gls{prm}
is defined as,
\begin{align}
  \text{PRM} &= \frac{P_{\text{firm}} - P_{\text{peak}}}{P_{\text{peak}}},
  \intertext{where}
  P_{\text{firm}} &= \text{installed firm capacity}, \nonumber\\
  P_{\text{peak}} &= \text{power demand at annual peak} \nonumber.
\end{align}
In other words, \gls{prm} is the fraction of capacity held in reserve during peak
times. ‘Affordability’ is an ill-defined term and varies significantly. An ``affordable''
energy system can be loosely understood as one with a low rate of energy poverty
\cite{brown_high_2020}. High energy costs have cascading adverse effects on every
aspect of life and significantly impact low-income households \cite{brown_high_2020}.
Measures that lead to increased costs for energy producers, such as carbon taxes,
almost always lead to higher consumer costs \cite{brown_high_2020,poelhekke_how_2019,khastar_how_2020}.
Thus, policymakers often seek a least-cost solution when designing energy policy.
Lastly, sustainability is defined as ``meeting the needs of the present without
compromising the ability of future generations to meet their own needs'' \cite{brook_why_2014,
the_united_nations_brundtland_commission_our_1987}. The twin problems of fossil
fuel consumption and climate change will significantly diminish the ability of
future generations to meet their basic needs. Droughts, forest fires, extreme
weather events, and flooding have increased due to climate change
\cite{reidmiller_fourth_2018}. Additionally, human activities like deforestation,
strip mining, and monocropping are causing a loss of essential ecosystem services
that advanced technology cannot easily replace \cite{malhi_climate_2020,butler_climate_2018,
costanza_value_1997}. These ecosystem services have an estimated total value between
\$125 trillion/year and \$145 trillion/year and, therefore, failure to mitigate
climate change will incur severe economic damage \cite{costanza_changes_2014, malhi_climate_2020}.
Climate change will affect each of these policy goals, which means minimizing its
effects is a moral, political, and economic imperative.

\subsection{Survey of climate mitigation policies}

There are three primary policy strategies for climate change mitigation \cite{fawzy_strategies_2020}:
conventional mitigation efforts, which include all policies centered around
reducing carbon emissions, negative emissions technologies, and the final option,
geoengineering, which reduces warming by increasing the Earth’s reflectivity. \gls{esom}
studies generally focus on the first set of energy policies, conventional mitigation strategies.
These strategies include:
\begin{itemize}
  \item Investing in clean energy resources, such as renewable energy and nuclear power
  \cite{fawzy_strategies_2020}. These strategies affect the topology of the energy
  system by influencing the types of energy production technologies.
  \item Demand response policies involve reducing grid load when there is insufficient
  supply forecasted or flattening the load profile throughout the day. These policies
  are critical for energy systems with high penetration of variable and intermittent
  energy sources like wind and solar power \cite{bouckaert_expanding_2014,
  kuzemko_policies_2017}. Demand response policies influence the temporal
  complexity of the energy system.
  \item Carbon capture and storage (CCS) reduces the carbon footprint of coal and
  natural gas plants but are not considered negative emissions \cite{fawzy_strategies_2020}.
  These strategies affect grid topology by enabling the continued use of previously
  high carbon-emitting technologies.
\end{itemize}

\subsection{Climate change strategies in Illinois}
\label{section:ceja}
In September of 2021, Illinois legislators passed a historic clean energy bill,
\gls{ceja} \cite{harmon_climate_2021}. This bill does several notable things for
both equity and energy policy. Naming just a few, \gls{ceja}:
\begin{enumerate}
  \item Sets Illinois on a path toward 100\% clean energy by 2050 and pledges 40\% renewable
  energy (e.g., wind or solar) by 2030 and 50\% by 2040
  \cite{office_of_governor_jb_pritzker_gov_2021}. Ambiguously,
  the text of the bill refers to 100\% renewable energy by 2050 \cite{harmon_climate_2021}.
  The former includes nuclear power, while the latter typically excludes nuclear power.
  \item Establishes a carbon credit plan that supports “zero emissions facilities”
  (i.e., nuclear power plants). Exelon reversed the imminent closure of Byron and
  Dresden nuclear power plants based on the financial support from this bill
  \cite{brown_two_2021}.
  \item Workforce development programs focused on equity. Such as a
  “Clean Jobs Workforce Network Hubs Program” that seeks to ensure “equity-focused”
  populations have support to enter the clean energy workforce
  \cite{office_of_governor_jb_pritzker_gov_2021}.
\end{enumerate}
This bill extends support for nuclear plants in the PJM Interconnect, where its
predecessor, the 2016 Future Energy Jobs Act, gave financial support to nuclear
plants in the MISO grid. The Solar Energy Industries Association (SEIA) and Illinois
Clean Jobs Coalition estimate that Illinois will have to add 6.3 GW to its wind capacity
and 17 GW of combined utility and residential solar energy
to achieve its renewable energy goals \cite{goeller_new_2021, the_accelerate_group_clean_2019}.
The Illinois Clean Jobs Coalition drove much of the content of the CEJA bill.
The Coalition funded a report to evaluate the economic impact of the CEJA bill,
finding that Illinois’ renewable energy goals could drive 39 billion dollars in
private investment to the state \cite{the_accelerate_group_clean_2019}.
However, the creators of this report did not consider the impacts of these targets
on grid operations, electricity cost, and energy reliability. Although nuclear power
plants (NPP) currently generate more than half of Illinois’ electricity, almost
90 percent of its carbon-free energy, politicians and lobbyists fiercely debate the continued
role of nuclear power in Illinois’ energy mix.

\subsection{Illinois Climate Action Plan at the University of Illinois}

The \gls{uiuc} released the first iteration
of the Illinois Climate Action Plan (iCAP) in 2010 and releases a new one every five years.
The 2015 iCAP report detailed UIUC’s goal to achieve carbon neutrality by 2050
\cite{institute_for_sustainability_energy_and_environment_illinois_2015}.
This goal includes every scope of emissions, from the energy produced on campus to
the emissions from vehicles faculty and staff use to drive to work
\cite{institute_for_sustainability_energy_and_environment_illinois_2015,
institute_for_sustainability_energy_and_environment_illinois_2020}. The
iCAP reports established several themes and objectives for each. These themes are:
\begin{itemize}
  \item Energy
  \item Transportation
  \item Land and Water
  \item Zero Waste
  \item Resilience
  \item Reporting Progress (accountability)
  \item Engagement
  \item Education
  \item Research Opportunities
  \item Funding
\end{itemize}
The ``Energy'' theme presents most significant opportunity to reduce carbon emissions
on the UIUC campus. \gls{uiuc}’s specific energy objectives include an energy
efficiency target of 167,000 BTU/square foot by 2030, purchasing 140 GWh of clean
electricity by 2025, and reducing building-level energy use
\cite{institute_for_sustainability_energy_and_environment_illinois_2020}. The University
heats virtually all of its buildings using steam produced by Abbott Power Plant
(APP), which burns coal and natural gas. The iCAP objectives conspicuously lack a
measurable target for clean thermal energy on campus.

\section{Energy System Optimization Models}

\glspl{esom} are a class of modeling tools designed to facilitate policy decisions before
future uncertainties are resolved \cite{hunter_modeling_2013}. In particular, \glspl{esom} answer questions
related to the planning and expansion of energy generating capacity \cite{de_queiroz_repurposing_2019}.
In order to achieve this goal, \glspl{esom} have to balance spatial, temporal, and topological
complexities while accounting for profound future uncertainties \cite{martinez-gordon_review_2021}.
\glspl{esom} typically use a “time-slice” approach where modelers use several
representative time slices to capture temporal variability \cite{fattahi_systemic_2020}.
Some frameworks handle spatial complexity by modeling different regions. The most detailed among
them incorporate data from geographical information systems (GIS) \cite{fattahi_systemic_2020}.
Lastly, most \glspl{esom} consider grid topology by grouping technology types (e.g. ‘coal power’
or ‘nuclear power’) \cite{fattahi_systemic_2020}, while some frameworks look at
interactions among individual plants \cite{jenkins_enhanced_2017}. Handling
future uncertainty is one of the most critical features of an \gls{esom}.

\subsection{Uncertainty Quantification}

Since perfect foresight is impossible, and policymakers often act then learn, handling
uncertainty is critical to developing robust insights
\cite{yue_review_2018, decarolis_modelling_2016}.
The two broad categories of uncertainties are parametric uncertainty and structural
uncertainty. Parametric uncertainty deals with the influence of unknown quantities
like future fuel prices, learning curves, and discount rates. Structural uncertainty
relates to factors not captured in a model’s equations\cite{hunter_modeling_2013,
yue_review_2018}. The latter uncertainty includes heterogeneity of decision-makers,
social acceptance, and different priorities \cite{yue_review_2018}. Most ESOM
studies handle uncertainties using a small ensemble of scenarios where a modeler
develops a  reference scenario and analyze policy impacts with additional constraints
and assumptions \cite{yue_review_2018}.  Several systematic methods account for
parametric uncertainty, including Monte Carlo, stochastic optimization, and
global sensitivity analysis \cite{yue_review_2018}. Structural uncertainty
is typically handled by either increasing model complexity to make it more
``realistic'' or studied systematically using \gls{mga}
\cite{hunter_modeling_2013,decarolis_modelling_2016, yue_review_2018}.
The framework I used in this thesis, \gls{temoa}, developed at North Carolina
State University, uses \gls{mga} to probe decision space efficiently
\cite{decarolis_temoa_2010}.


\section{Survey of \gls{esom} Literature}

\glspl{esom} provide valuable policy insights to decision-makers. \gls{temoa}
is one modeling framework among many and is structurally similar to TIMES/MARKAL
and MESSAGE since they are formulated with linear programming \cite{yue_review_2018}.
OpenModelica models technologies with much greater technical detail than traditional
\glspl{esom} and are closer to unit-commitment models. I include them in this review
because they use OpenModelica and the \gls{raven} tool to study \glspl{nhes}
\cite{baker_optimal_2018,garcia_dynamic_2016,epiney_economic_2020}.
The tradeoff between model complexity and computational cost is well documented.
Every study described in Table \ref{tab:esom_lit} with at least an hourly
time resolution optimizes at most a single year. For example, Garcia et al. 2016,
Epiney et al. 2020, and Baker et al. 2018 all use hourly resolution, capturing
the variability of solar and wind. However, they only consider single microgrids
and optimize a single year
\cite{epiney_economic_2020,baker_optimal_2018,garcia_dynamic_2016}.
These models do not consider demand growth, existing capacity, or energy policy
scenarios. Komiyama et al. 2015 used a 5-minute time resolution for a detailed
solar and wind variability model. That work only optimizes one year, but
the study does consider several emissions and energy policies \cite{komiyama_energy_2015}.
The most spatially complex model, Neumann\&Brown 2021, consider 100 spatial nodes
spanning the United Kingdom and continental Europe along with 12 times of day for
an entire year (4380 time slices) \cite{neumann_near-optimal_2021}. Neumann\&Brown
further use \gls{mga} to examine near-optimal solutions. However, Neumann\&Brown consider
a limited number of technology options and only a single energy sector, electricity.
Other works employ various dimensionality reduction techniques, such as matching
load curves \cite{de_sisternes_value_2016,poncelet_impact_2016} or choosing time
slices using k-means \cite{kotzur_impact_2018} to model temporal variations efficiently.
However, every work examining multidecadal time horizons relies on a limited number
of time slices to model each year
\cite{alzbutas_uncertainty_2012,barron_differential_2015,bennett_extending_2021,
bouckaert_expanding_2014,decarolis_modelling_2016,li_open_2020,poncelet_impact_2016,
seck_embedding_2020,yue_least_2020}. These models may use as few as four annual
time slices \cite{decarolis_modelling_2016} or up to 288 annual time slices \cite{poncelet_impact_2016}.
Here, the primary tradeoff is between temporal variations and topological complexity.
Poncelet et al. 2016 recommend prioritizing temporal complexity over topological
complexity due to the strong impact of resource variability on model results \cite{poncelet_impact_2016}.

The sensitivity analysis in the State of Illinois case study explicitly
examines the effect of annual variability wind and solar resources by varying
the capacity factor of those technologies. Such analysis is absent from the literature
but is essential to answering the fundamental question of \glspl{esom}: How much
capacity is enough?

  \begin{table}
    \centering
    \caption{Summary of ESOM Literature Survey}
    \label{tab:esom_lit}
    % \input{tables/esom_lit_review}
    \resizebox{\textwidth}{!}{\input{tables/esom_lit_review}}
  \end{table}


% \newgeometry{margin=1cm}
% \begin{landscape}
%   \begin{table}
%     \centering
%     \caption{Summary of ESOM Literature Survey}
%     \label{tab:esom_lit}
%     % \input{tables/esom_lit_review}
%     \resizebox{1.25\textwidth}{!}{\input{tables/esom_lit_review}}
%   \end{table}
% \end{landscape}
% \restoregeometry
