The literature review in chapter \ref{chapter:litrev} showed the tradeoffs among
different types of model complexity. Although some studies considered temporal
resolution, which captures intra-year variability, there were no studies found in
the literature review that consider the influence of \textit{inter}-year variability,
due in part to \glspl{esom}' assumption that all modeled years are identical.
This chapter introduces the software and tools used to conduct experiments in this
thesis that explicitly examine the effect of annual variability of wind and solar
resources. As well as describing the case studies and experiments themselves.
The tools I used were:
\begin{itemize}
  \item Tools for Energy System Optimization and Analysis (\texttt{Temoa}), and
  \item Python for Generating Energy Systems (\pygen).
\end{itemize}
In this thesis, I demonstrate the role of advanced nuclear reactors to meet future
energy demands with two case studies: Illinois’ electric grid and the microgrid
at the University of Illinois at Urbana-Champaign. Both case studies use \gls{temoa}
\cite{decarolis_temoa_2010} to solve for the optimal least-cost solution. I
developed \gls{pygen} \cite{dotson_python_2021} to facilitate the creation of
\gls{temoa} input files and sensitivity analyses.

\section{Tools for Energy System Optimization and Analysis (\texttt{Temoa})}
\input{methods/temoa}

\section{Python for Generating Energy Systems (\texttt{PyGenesys})}
\input{methods/pygen}


\section{Case Study I: Carbon Zero Electricity in Illinois}
\label{section:ilmodel}
\input{methods/illinois_case_desc}

\section{Case Study II: The UIUC Microgrid}
\label{section:uiucmodel}
\input{methods/uiuc_case_desc.tex}

\section{What About Geothermal, Biomass, and Hydropower?}
\label{section:alt_firm}

Nuclear energy has significant barriers to acceptance, such as nuclear waste,
accident risk, and fear of weapons proliferation. Whether or not these barriers
constitute genuine technological challenges is beside the point; some decision-makers
may look for alternatives to nuclear energy that can serve similar roles
on the grid. Geography constrains hydropower and geothermal power. \gls{nrel} analyzed
the technical potential for each renewable resource and found that hydroelectric dams
and conventional geothermal plants are negligible options in Illinois. Even run-of-the-river
hydropower could contribute less than a gigawatt \cite{lopez_us_2012}. Enhanced Geothermal
Systems (EGS) can overcome geographic challenges by creating artificial geological permeability
with a method similar to hydraulic fracturing \cite{geothermal_technologies_office_what_2012}.
However, EGSs have high capital costs, more than twice an AP1000 Nuclear Reactor
\cite{nrel_2020_2020}. \gls{uiuc} has a geothermal loop that supplies 135 tons of refrigeration
and heating, up to 65\% of the energy use for one new (as of 2021) building on campus,
the Campus Instructional Facility \cite{kim_geothermal_nodate}.
This loop cost \gls{uiuc} nearly 2 million dollars which is 580\% more expensive
than the existing chilled water capacity
\cite{kim_geothermal_nodate,affiliated_engineers_inc_utilities_2015}. For these
reasons, I did not model geothermal or hydropower for either \gls{uiuc} or Illinois.
Dedicated biomass power plants are modeled for Illinois to offer an alternative
firm capacity option. However, \gls{nrel} also modeled the technical potential for biopower
and found that Illinois could generate up to 4 GW of electricity from biofueled turbines, the highest potential
for any U.S. state \cite{lopez_us_2012}. A further caveat to this technical potential is that it does not
consider alternative uses for the land (e.g., food production). Nevertheless, a 4 GW limit is a constraint in the Illinois study for the time resolution analysis.
Table \ref{tab:nuclear_alt} summarizes important parameters for firm capacity options.

\begin{table}[H]
  \centering
  \caption{Summary of Firm Capacity Options}
  \label{tab:nuclear_alt}
\resizebox{\textwidth}{!}{  \begin{tabular}{lrrrrr}
    \toprule
    Technology & Capacity Factor & Technical Potential$^\text{a}$ & Capital Cost &
    Lifecycle Emissions\\
     & [\%] & [GWe] & [M\$/GW]& g CO$_2$$_{eq}$/kWh \\
     Source & \cite{energy_information_administration_capacity_2020,nrel_2020_2020}&
     \cite{lopez_us_2012}& \cite{nrel_2020_2020} & \cite{nicholson_life_2021}\\
    \midrule
    Nuclear & 92.4 & - & 6.1 & 13 \\
    Biomass & 62.5 & 4 & 4.0 & 42 \\
    Geothermal$^\text{b}$ & 69.1 & 0 & 4.5 & 29\\
    Enhanced Geothermal$^\text{b}$ & 90.0 & 86 & 14.5 & 32 \\
    Hydroelectric & 40.7 & 1 & 3.9-6.2 & 21 \\
    \bottomrule
    \multicolumn{5}{l}{$^\text{a}$ Assumes all energy and fuel is produced in Illinois}\\
    \multicolumn{5}{l}{$^\text{b}$ Assumes ``flash'' technology}
  \end{tabular}}
\end{table}
